Attract and kill: Spinosad containing spheres to control onion maggot (\textit{Delia antiqua})

### 2020-01-20 manuscript/pest-management-science/reviewer-response/reviewer-response.tex

The authors should add a few sentences describing the field sites (lines 54-56) cropping history (e.g. Where onions grown in previous years?) and the reason why the fields were chosen (e.g. did the fields have a previous history of onion maggot damage?).  If experiments were not replicated in space or time it should be noted in the Material and Methods.

\begin{quote}
    \textit{Thanks for mentioning this opportunity to elaborate.  Details on cropping history and reasoning behind field site choice have now been included on lines ....  Information on how experiments were replicated in space and time are also included (see above).  }
\end{quote}

In the discussion section the authors made no mention of using spinosad seed treatment as a management option in conventional and organic onion fields.  Why?? Spinosad (Regard) seed treatment is currently the primary management method for onion maggot in organic onions, and spinosad is an active ingredient in many conventional seed treatments.   Spinosad is an active ingredient in the FI500 seed treatment, and Regard alone often provides a similar level of protection compared to FI500.  Since FI500 resulted in lower maggot damage compared to all "attract and kill" tactics, why would someone recommend attract and kill with spinosad over spinosad seed treatment?   I strongly recommend the authors discuss the advantages, disadvantages, and compatibility of using spinosad seed treatment and spinosad attract & kill.

\begin{quote}
    \textit{This is an excellent point.  We've brought this up and elaborated more on exactly this point in the discussion as well as addressing the advantages and disadvantages of these strategies.  Thanks for the suggestion.  Additions can be found on lines ....
    }
\end{quote}

@@

The authors should discuss the implications of onion maggot developing resistance to spinosad using the attract and kill tactic.  e.g. Will the attract and kill tactic with spinosad lead to a higher likelihood of onion maggot developing resistance to spinosad?  This tactic targets egg laying adults, it only provides partial control, and it exposes multiple generations per year to spinosad?   These factors seem to favor resistance. This question is especially relevant as the most popular registered insecticide seed treatments contain spinosad and there are few alternatives in the marketplace.

\begin{quote}
    \textit{Also an excellent suggestion.  This is particularly important in systems where options are rapidly disappearing.  We've addressed this in more detail in the discussion on lines ...
    }
\end{quote}

@@

Line 112: aren't these the same thing?

\begin{quote}
    \textit{Thanks for your question.  Yes and No.  While linear models and analysis of variance use the same underlying model structure, they have two different objectives and answer two different questions.  Linear models represent the relationship between independent and dependent variables.  Analysis of variance compares inter and intra-group variation to examine differences.  In our case, both approaches were used because we are interested both in modeling relationships and differences between groups.  }
\end{quote}

Line 113: I'd like to get more details about which factors you have used, fixed and random.

\begin{quote}
    \textit{Thanks for suggesting this inclusion.  We've listed the fixed and random factors used in the analysis on lines ...}
\end{quote}

@@

Line 115: AIC?, BIC? How did you use these to inform model selection? Considerations: third time this word is used in this sentence.

\begin{quote}
    \textit{Good question.  Both were used to avoid overfitting.  As both of these criteria penalize likelihood estimates with increasing factors, these were used in consideration with other factors listed to determine best fit models.  Redundant uses of the word 'considerations' have been removed from the sentence.  
    }
\end{quote}

@@

Line 116:  I suggest rewording this sentence. I'm assuming that you are talking about means separation here?

\begin{quote}
    \textit{Exactly!  We've reworded this sentence to clarify.  Thanks for the suggestion.  }
\end{quote}

Line 116: Why use both, why not just one or the other?

\begin{quote}
    \textit{Tukey HSD test generally is used to evaluate all pairwise comparisons.  Dunnett's test evaluates treatments versus controls.  In some cases we are only interested in comparisons versus controls (more powerful statistically as the correction for the family wise error rate is lower for less comparisons) while in others we are interested in more detailed comparisons.  }
\end{quote}

Line 117: Be more explicit about which model goes with what data. I'd like to see a list of dependent and independent variables and fixed and random effects.I suggest removing such a blanket statement and being more specific as I mentioned above.

\begin{quote}
    \textit{Thanks for the suggestion.  We've included more detail about the models and factors used for each datasets on lines ...  In addition, the exact code to run the analysis with annotations and extensive documentation is publicly available on GitHub.  }
\end{quote}

@@

Line 124: need to mention these where you talk about how you built and selected models. These are the tools you used to do that.

\begin{quote}
    \textit{Thanks for the suggestion.  Given that these tools were used throughout the analysis, we thought it best to consolidate them here rather than repeatedly mentioning them throughout the text.  }
\end{quote}

@@

Line 128: I'm not following why showing this data is relevant. You also have not done analysis on it. Suggest incorporating it into the rest of the MS effectively.

\begin{quote}
    \textit{Thanks for the opportunity to elaborate.  These data are important as they document three generations in upstate New York.  This is crucial because it highlights the importance of targeting the first generation to prevent further damage throughout the season.  While the these data are often mentioned, this figure is important to practitioners in New York state for reference and management decisions.  Generalized additive models with cubic splines were used to highlight trends in data for management professionals.  }
\end{quote}

@@

Line 129: What period was this? Be more specific

\begin{quote}
    \textit{Excellent opportunity to clarify.  Thanks.  We've added details to lines ...}
\end{quote}

@@

Line 131: delete "roughly"

\begin{quote}
    \textit{Excellent suggestion.  'Roughly' has been deleted.  }
\end{quote}

Line 133: What do you mean? Significantly different? Different from what?

\begin{quote}
    \textit{Great opportunity to clarify.  This has been explained on lines...}
\end{quote}

Line 143: Was this peak significantly different from the other times?

\begin{quote}
    \textit{Interesting question.  While we did not specifically test for this, peaks were either markedly reduced or noticeably absent for subsequent times in Figure 3.  }
\end{quote}

Line 154: This is compared to what?

\begin{quote}
    \textit{This comparison was made to spinosad spheres alone - specifics have been addded to lines ...}
\end{quote}

Line 155: I have a hard time with the letters above bars on this figure. I didn't see the darker colored 'A's. Usually 'A' is used for the highest numbers.

\begin{quote}
    \textit{Thanks for this suggestion.  We've updated the colors to make this darker.  The lettering scheme was chosen to read alphabetically from left to right.  }
\end{quote}

Line 155: Was this significant?

\begin{quote}
    \textit{Great question.  It was indeed significant.  Details have now been included in lines ...}
\end{quote}

Line 157: be more specific about what treatments.

\begin{quote}
    \textit{In this case treatment was with the attract and kill strategies.  Details have been added to lines ...}
\end{quote}

Line 159: are there any F-values or model results to go along with this?

\begin{quote}
    \textit{Thanks for the opportunity to add more detail.  Additional model statistics have been included on lines ...}
\end{quote}

\textbf{
Line 162: The discussion and introduction seems to have been written by different people. The language used in the discussion is much more "fuzzy" than in the intro. I suggest re-wording much of the discussion to make it less wordy and repetitive and place the findings into a larger context.
}

Line 167: was this personal observation or speculation?

Line 169: Suggest simplifying this paragraph, its a long and convoluted way of saying that dose didn't matter.

### 2020-01-20 manuscript/pest-management-science/main.tex
Updated reviewer responses

To evaluate adult \textit{D. antiqua} mortality as a result of contact with spheres containing spinosad, linear models and analysis of variance were used.  Models were determined after consideration of all factor combinations and interactions and selected based on consideration of residual diagnostics (conformance to assumptions of normality and homoscedasticity), goodness of fit tests, $R^2$ values, information criteria, and leverage considerations. Main effects were teased out from interactions using Tukey and Dunnett post-hoc tests at $P < 0.05$.

@@

During the period these trials were conducted in upstate New York, \textit{D. antiqua} had three generations (Figure \ref{fig:figure1}).  The first population of adults peaked in June, the second in late July, and the third in August.  More males than females were collected on sticky cards, but both males and females followed roughly the same population cycles.

@@

White spheres containing spinosad caused appreciable mortality of \textit{D. antiqua} adults (Figure \ref{fig:figure2}A).  Both spinosad rate (F =114.3, df=2, p \textless 0.001)  and exposure time (F = 4.7, df = 1, p = 0.03) in the field significantly explained approximately 62\% of the observed variation in adult \textit{D. antiqua} mortality ($R^2_{adj}$ = 0.62, p \textless 0.001).  Both rates of spinosad treated spheres caused approximately 47$\pm$4\% more \textit{D. antiqua} mortaliy than untreated controls (Figure \ref{fig:figure2}B).  Increasing rates of spinosad from 0.5\% to 1\% did not signifcantly increase \textit{D. antiqua} mortality (Figure \ref{fig:figure2}B).

@@

Deployment of the attract and kill solution of spinosad containing spheres did influence cumulative onion damage in the field as a result of \textit{D. antiqua} larval feeding (Figure \ref{fig:figure4}A).  Cumulative damage as a result of \textit{D. antiqua} larval feeding was significantly explained by treatment solution ($\chi^2$ = 42.2, df = 3, p \textless 0.001), the presence of Delia Lure ($\chi^2$ = 10.0, df = 1, p = 0.001), and their interaction ($\chi^2$ = 21.4, df = 1 p \textless 0.001).  Cumulative damage due to feeding by \textit{D. antiqua} larvae over the course of the field season ranged from 0 in some control plots to 82 dead \textit{D. antiqua} plants in a few plots. This amounts to approximately 20-30\% losses of onion directly attributable feeding by \textit{D. antiqua} larvae.  All treatments were significantly worse than insecticide controls which had very low levels of \textit{D. antiqua} damage (Figure \ref{fig:figure4}A).  Pairing Delia Lure with spinosad containing spheres significantly reduced damage by \textit{D. antiqua} larvae by 72$\pm$10\% (p \textless 0.001) compared to spinosad spheres alone.  Pairing Delia Lure with sticky cards had the opposite effect (Figure \ref{fig:figure4}A).

@@

Treatment significantly explained observed \textit{D. antiqua} damage in field cage trials with the enhanced formulation ($\chi^2$ = 23.9, df = 2, p \textless 0.001).  The enhanced formulation (spinosad +) containing 0.5\% spinosad, 5.0\% ammonium carbonate and 5.0\% casein hydrolysate did not perform significantly different than spinosad-only treatments with 0.5\% spinosad containing spheres (Figure \ref{fig:figure4}B).  Both treatments containing spinosad performed significantly better than the no insecticide controls (p \textless 0.05, Figure \ref{fig:figure4}B).

@@

Using \textit{D. antiqua} population averages from field experiments and mortality achieved by field-exposed spheres over time assessed in the laboratory, numbers of adult \textit{D. antiqua}  killed as a result of the attract and kill strategy can be estimated. During peak flights, approximately 272 \textit{D. antiqua} adults (both males and females) were caught on average per week.  Over the course of the field season, approximately 64 adults were caught by sticky cards on average per week.  Assuming spinosad containing spheres kill approximately 54\% of the adults they contact, over the course of a 16 week season, this attract and kill solution could have killed approximately 553 \textit{D. antiqua} adults (16 weeks x 64 adults on average per week * 0.54 mortality) per spinosad containing sphere.

>>>>

To evaluate adult \textit{D. antiqua} mortality as a result of contact with spheres containing spinosad, linear models and analysis of variance were used.  Models were determined after consideration of all factor combinations (rate and exposure time for mortality trials and control strategy and presence of Delia Lure with date as a random effect for field trials) and interactions and selected based on residual diagnostics (conformance to assumptions of normality and homoscedasticity), goodness of fit tests, $R^2$ values, information criteria (AIC and BIC), leverage, and likelihood ratio tests. Main effects and means separations were evaluated using Tukey and Dunnett post-hoc tests at $P < 0.05$.

@@

Over the course of the 2006 field season when population surveys were conducted in upstate New York, \textit{D. antiqua} had three generations (Figure \ref{fig:figure1}).  The first population of adults peaked in June, the second in late July, and the third in August.  More males than females were collected on sticky cards, but both males and females followed roughly the same population cycles.

@@

White spheres containing spinosad caused significant mortality of \textit{D. antiqua} adults compared to controls (Figure \ref{fig:figure2}A).  Both spinosad rate (F =114.3, df=2, p \textless 0.001)  and exposure time (F = 4.7, df = 1, p = 0.03) in the field significantly explained approximately 62\% of the observed variation in adult \textit{D. antiqua} mortality ($R^2_{adj}$ = 0.62, p \textless 0.001).  Both rates of spinosad treated spheres caused approximately 47$\pm$4\% more \textit{D. antiqua} mortaliy than untreated controls (Figure \ref{fig:figure2}B).  Increasing rates of spinosad from 0.5\% to 1\% did not signifcantly increase \textit{D. antiqua} mortality (Figure \ref{fig:figure2}B).

@@

Deployment of the attract and kill solution of spinosad containing spheres did influence cumulative onion damage in the field as a result of \textit{D. antiqua} larval feeding (Figure \ref{fig:figure4}A).  Cumulative damage as a result of \textit{D. antiqua} larval feeding was significantly explained by treatment solution ($\chi^2$ = 42.2, df = 3, p \textless 0.001), the presence of Delia Lure ($\chi^2$ = 10.0, df = 1, p = 0.001), and their interaction ($\chi^2$ = 21.4, df = 1 p \textless 0.001).  Cumulative damage due to feeding by \textit{D. antiqua} larvae over the course of the field season ranged from 0 in some control plots to 82 dead \textit{D. antiqua} plants in a few plots. This amounts to approximately 20-30\% losses of onion directly attributable feeding by \textit{D. antiqua} larvae.  All treatments were significantly worse than insecticide controls which had very low levels of \textit{D. antiqua} damage (Figure \ref{fig:figure4}A).  Pairing Delia Lure with spinosad containing spheres significantly reduced damage by \textit{D. antiqua} larvae by 72$\pm$10\% (p \textless 0.001) compared to spinosad spheres alone.  Pairing Delia Lure with sticky cards had the opposite effect (P = 0.002, Figure \ref{fig:figure4}A).

@@

Treatment (implementation of attract and kill strategies) significantly explained observed \textit{D. antiqua} damage in field cage trials with the enhanced formulation ($\chi^2$ = 23.9, df = 2, p \textless 0.001).  The enhanced formulation (spinosad +) containing 0.5\% spinosad, 5.0\% ammonium carbonate and 5.0\% casein hydrolysate did not perform significantly different (t = -1.9, df = 64, p = 0.13) than spinosad-only treatments with 0.5\% spinosad containing spheres (Figure \ref{fig:figure4}B).  Both treatments containing spinosad performed significantly better than the no insecticide controls (Spinosad: t = 4.9, df = 64, p \textless 0.0001; Spinosad +: t = 2.9, df = 64, p = 0.01, Figure \ref{fig:figure4}B).

@@

The viability of low doses could be particularly important in light of other uses of spinosad in onion fields.  Spinosad seed treatment is a widespread method used for control of onion maggot and common in other commercial seed treatments \citep{nault2006performance,nault2006onion, wilson2015evaluation}.  It is also present in the FI500 seed treatment and alone can offer comparable control \citep{nault2006performance,nault2006onion, wilson2015evaluation}.  These seed treatments are effective options of control of the larval stage; attract and kill methodologies using spinosad evaluated here target the mobile adults.  While exposure to an additional source of spinosad may foster development of resistance, strategies for preventing development of resistance could include rotating use of the active ingredient for the 'kill' part of the attract and kill strategy.

Using \textit{D. antiqua} population averages from field experiments and mortality achieved by field-exposed spheres over time assessed in the laboratory, numbers of adult \textit{D. antiqua}  killed as a result of the attract and kill strategy using spinosad can be estimated. During peak flights, approximately 272 \textit{D. antiqua} adults (both males and females) were caught on average per week.  Over the course of the field season, approximately 64 adults were caught by sticky cards on average per week.  Assuming spinosad containing spheres kill approximately 54\% of the adults they contact, over the course of a 16 week season, this attract and kill solution could have killed approximately 553 \textit{D. antiqua} adults (16 weeks x 64 adults on average per week * 0.54 mortality) per spinosad containing sphere.
